---
title: "Tutorial"
tutorial:
  id: "debc2ab2-a03e-4b1f-ba1f-72dd21fac050"
  version: 3.1
output: 
    learnr::tutorial:
        progressive: true
        allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(checkr)
knitr::opts_chunk$set(echo = FALSE)
tutorial_options(exercise.checker = checkr::check_for_learnr)
```

## Topic 0

### My Exercise

Use the rep() function, with 1:4 as an input, to generate the 12-element vector 1, 1, 1, 2, 2, 2, 3, 3, 3, 4, 4, 4

```{r exercise1, exercise=TRUE}

```


```{r exercise1-check}
check_exer_1_v0(USER_CODE)
```


## Topic 1

### Exercise 

*Here's a simple exercise with an empty code chunk provided for entering the answer.*

Write the R code required to add two plus two:

```{r add-two-and-two, exercise=TRUE, exercise.lines=1}


```

```{r add-two-and-two-check}
check_add_two_and_two(USER_CODE)
```

### Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaulated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

## Topic 3

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

