---
title: "Chapter 01 of DSUR"
author: "Peter Baumgartner"
date: "`r Sys.Date()`"
output: 
    learnr::tutorial:
        version: 0.1
        progressive: true
        allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(checkr)
knitr::opts_chunk$set(echo = FALSE)
tutorial_options(exercise.checker = checkr::check_for_learnr)
# assign("my_points", 0, envir = .GlobalEnv)
```

# DSUR: Chapter 01

<!-- ## Quiz -->

<!-- ### Question 01 -->

<!-- ```{r quiz-chapter-01-01} -->
<!--   question("What is a confounding variable?", -->
<!--     answer("A variable that is manipulated by the experimenter."), -->
<!--     answer("A variable that affects the outcome being measured as well as or instead of the independent variable.", correct = TRUE), -->
<!--     answer("A variable that has not been measured."), -->
<!--     answer("A variable that is made up only of categories."), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!--   ) -->
<!-- ``` -->

<!-- ### Question 02 -->


<!-- ```{r quiz-chapter-01-02} -->
<!-- question("'Children can learn a second language faster before the age of 7'. Is this statement:", -->
<!--     answer("A null hypothesis"), -->
<!--     answer("A non‐scientific statement."), -->
<!--     answer("A two‐tailed hypothesis."), -->
<!--     answer("A one‐tailed hypothesis.", correct = TRUE), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!--   ) -->
<!-- ``` -->

<!-- ### Question 03 -->

<!-- ```{r quiz-chapter-01-03} -->
<!-- question("If a psychological test is valid, what does this mean?", -->
<!--     answer("The test will give consistent results."), -->
<!--     answer("The test measures what it claims to measure.", correct = TRUE), -->
<!--     answer("The test has internal consistency."), -->
<!--     answer("The test measures a psychologically useful variable."), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 04 -->

<!-- ```{r quiz-chapter-01-04} -->
<!-- question("If my null hypothesis is ‘Dutch people do not differ from English people in height’, what is my alternative hypothesis?", -->
<!--     answer("Dutch people are taller than English people."), -->
<!--     answer("English people are taller than Dutch people."), -->
<!--     answer("Dutch people differ in height from English people."), -->
<!--     answer("All of the above are plausible alternative hypotheses.", correct = TRUE), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 05 -->

<!-- ```{r quiz-chapter-01-05} -->
<!-- question("When questionnaire scores predict, or correspond with, external measures of the same construct that the questionnaire measures it is said to have:", -->
<!--     answer("Ecological validity."), -->
<!--     answer("Factorial validity."), -->
<!--     answer("Content validity."), -->
<!--     answer("Criterion validity.", correct = TRUE), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 06 -->

<!-- ```{r quiz-chapter-01-06} -->
<!-- question("A variable manipulated by a researcher is known as:", -->
<!--     answer("A dependent variable."), -->
<!--     answer("A confounding variable."), -->
<!--     answer("A discrete variable."), -->
<!--     answer("An independent variable.", correct = TRUE), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 07 -->

<!-- ```{r quiz-chapter-01-07} -->
<!-- question("A predictor variable is another name for:", -->
<!--     answer("A dependent variable."), -->
<!--     answer("A confounding variable."), -->
<!--     answer("A discrete variable."), -->
<!--     answer("An independent variable.", correct = TRUE), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 08 -->

<!-- ```{r quiz-chapter-01-08} -->
<!-- question("What kind of variable is IQ, measured by a standard IQ test?", -->
<!--     answer("Categorical."), -->
<!--     answer("Discrete."), -->
<!--     answer("Nominal."), -->
<!--     answer("Continuous.", correct = TRUE), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 09 -->

<!-- ```{r quiz-chapter-01-09} -->
<!-- question("A frequency distribution in which high scores are most frequent (i.e. bars on the graph are highest on the right‐hand side) is said to be:", -->
<!--     answer("Positively skewed."), -->
<!--     answer("Leptokurtic."), -->
<!--     answer("Platykurtic."), -->
<!--     answer("Negatively skewed.", correct = TRUE), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 10 -->

<!-- ```{r quiz-chapter-01-10} -->
<!-- question("A frequency distribution in which there are too few scores at the extremes of the distribution is said to be:", -->
<!--     answer("Positively skewed."), -->
<!--     answer("Leptokurtic.", correct = TRUE), -->
<!--     answer("Platykurtic."), -->
<!--     answer("Negatively skewed."), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 11 -->

<!-- ```{r quiz-chapter-01-11} -->
<!-- question("Which of the following is designed to compensate for practice effects?", -->
<!--     answer("A repeated measures design."), -->
<!--     answer("Randomization of participants."), -->
<!--     answer("Counterbalancing.", correct = TRUE), -->
<!--     answer("A control condition."), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 12 -->

<!-- ```{r quiz-chapter-01-12} -->
<!-- question("Variation due to variables that have not been measured is known as:", -->
<!--     answer("Unsystematic variance.", correct = TRUE), -->
<!--     answer("Homogeneous variance."), -->
<!--     answer("Systematic variance."), -->
<!--     answer("Model variance."), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 13 -->

<!-- ```{r quiz-chapter-01-13} -->
<!-- question("If the scores on a test have a mean of 26 and a standard deviation of 4, what is the z‐score for a score of 18?", -->
<!--     answer("-2", correct = TRUE), -->
<!--     answer("11"), -->
<!--     answer("2"), -->
<!--     answer("-1.41"), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

## Second Quiz

### 13 Questions


```{r quiz-chapter-01}
my_good_answers <- 0
my_bad_answers <- 0

quiz(caption = "Quiz Chapter 01",
  question("What is a confounding variable?",
    answer("A variable that is manipulated by the experimenter."),
    answer("A variable that affects the outcome being measured as well as or instead of the independent variable.", correct = TRUE),
    answer("A variable that has not been measured."),
    answer("A variable that is made up only of categories."),
    random_answer_order = TRUE,
    correct = paste("Correct: You've reached", my_good_answers <- my_good_answers + 1, " of 13 possible points."),
    incorrect = paste("Incorrect: You've reached", my_bad_answers <- my_bad_answers + 1, " of 13 possible points.")
  ),
  
question("'Children can learn a second language faster before the age of 7'. Is this statement:",
    answer("A null hypothesis"),
    answer("A non‐scientific statement."),
    answer("A two‐tailed hypothesis."),
    answer("A one‐tailed hypothesis.", correct = TRUE),
    random_answer_order = TRUE,
    correct = paste("Correct: You've reached", my_good_answers <- my_good_answers + 1, " of 13 possible points."),
    incorrect = paste("Incorrect: You've reached", my_bad_answers <- my_bad_answers + 1, " of 13 possible points.")
  ),

question("If a psychological test is valid, what does this mean?",
    answer("The test will give consistent results."),
    answer("The test measures what it claims to measure.", correct = TRUE),
    answer("The test has internal consistency."),
    answer("The test measures a psychologically useful variable."),
    random_answer_order = TRUE,
    correct = paste("Correct: You've reached", my_good_answers <- my_good_answers + 1, " of 13 possible points."),
    incorrect = paste("Incorrect: You've reached", my_bad_answers <- my_bad_answers + 1, " of 13 possible points.")
),

question("If my null hypothesis is ‘Dutch people do not differ from English people in height’, what is my alternative hypothesis?",
    answer("Dutch people are taller than English people."),
    answer("English people are taller than Dutch people."),
    answer("Dutch people differ in height from English people."),
    answer("All of the above are plausible alternative hypotheses.", correct = TRUE),
    random_answer_order = TRUE,
    correct = paste("Correct: You've reached", my_good_answers <- my_good_answers + 1, " of 13 possible points."),
    incorrect = paste("Incorrect: You've reached", my_bad_answers <- my_bad_answers + 1, " of 13 possible points.")
),

question("When questionnaire scores predict, or correspond with, external measures of the same construct that the questionnaire measures it is said to have:",
    answer("Ecological validity."),
    answer("Factorial validity."),
    answer("Content validity."),
    answer("Criterion validity.", correct = TRUE),
    random_answer_order = TRUE,
    correct = paste("Correct: You've reached", my_good_answers <- my_good_answers + 1, " of 13 possible points."),
    incorrect = paste("Incorrect: You've reached", my_bad_answers <- my_bad_answers + 1, " of 13 possible points.")
),

question("A variable manipulated by a researcher is known as:",
    answer("A dependent variable."),
    answer("A confounding variable."),
    answer("A discrete variable."),
    answer("An independent variable.", correct = TRUE),
    random_answer_order = TRUE,
    correct = paste("Correct: You've reached", my_good_answers <- my_good_answers + 1, " of 13 possible points."),
    incorrect = paste("Incorrect: You've reached", my_bad_answers <- my_bad_answers + 1, " of 13 possible points.")
),

question("A predictor variable is another name for:",
    answer("A dependent variable."),
    answer("A confounding variable."),
    answer("A discrete variable."),
    answer("An independent variable.", correct = TRUE),
    random_answer_order = TRUE,
    correct = paste("Correct: You've reached", my_good_answers <- my_good_answers + 1, " of 13 possible points."),
    incorrect = paste("Incorrect: You've reached", my_bad_answers <- my_bad_answers + 1, " of 13 possible points.")
),

question("What kind of variable is IQ, measured by a standard IQ test?",
    answer("Categorical."),
    answer("Discrete."),
    answer("Nominal."),
    answer("Continuous.", correct = TRUE),
    random_answer_order = TRUE,
    correct = paste("Correct: You've reached", my_good_answers <- my_good_answers + 1, " of 13 possible points."),
    incorrect = paste("Incorrect: You've reached", my_bad_answers <- my_bad_answers + 1, " of 13 possible points.")
),

question("A frequency distribution in which high scores are most frequent (i.e. bars on the graph are highest on the right‐hand side) is said to be:",
    answer("Positively skewed."),
    answer("Leptokurtic."),
    answer("Platykurtic."),
    answer("Negatively skewed.", correct = TRUE),
    random_answer_order = TRUE,
    correct = paste("Correct: You've reached", my_good_answers <- my_good_answers + 1, " of 13 possible points."),
    incorrect = paste("Incorrect: You've reached", my_bad_answers <- my_bad_answers + 1, " of 13 possible points.")
),

question("A frequency distribution in which there are too few scores at the extremes of the distribution is said to be:",
    answer("Positively skewed."),
    answer("Leptokurtic.", correct = TRUE),
    answer("Platykurtic."),
    answer("Negatively skewed."),
    random_answer_order = TRUE,
    correct = paste("Correct: You've reached", my_good_answers <- my_good_answers + 1, " of 13 possible points."),
    incorrect = paste("Incorrect: You've reached", my_bad_answers <- my_bad_answers + 1, " of 13 possible points.")
),

question("Which of the following is designed to compensate for practice effects?",
    answer("A repeated measures design."),
    answer("Randomization of participants."),
    answer("Counterbalancing.", correct = TRUE),
    answer("A control condition."),
    random_answer_order = TRUE,
    correct = paste("Correct: You've reached", my_good_answers <- my_good_answers + 1, " of 13 possible points."),
    incorrect = paste("Incorrect: You've reached", my_bad_answers <- my_bad_answers + 1, " of 13 possible points.")
),

question("Variation due to variables that have not been measured is known as:",
    answer("Unsystematic variance.", correct = TRUE),
    answer("Homogeneous variance."),
    answer("Systematic variance."),
    answer("Model variance."),
    random_answer_order = TRUE,
    correct = paste("Correct: You've reached", my_good_answers <- my_good_answers + 1, " of 13 possible points."),
    incorrect = paste("Incorrect: You've reached", my_bad_answers <- my_bad_answers + 1, " of 13 possible points.")
),

question("If the scores on a test have a mean of 26 and a standard deviation of 4, what is the z‐score for a score of 18?",
    answer("-2", correct = TRUE),
    answer("11"),
    answer("2"),
    answer("-1.41"),
    random_answer_order = TRUE,
    correct = paste("Correct: You've reached", my_good_answers <- my_good_answers + 1, " of 13 possible points."),
    incorrect = paste("Incorrect: You've reached", my_bad_answers <- my_bad_answers + 1, " of 13 possible points.")
    )
)
```


## Basics: Measures of Central Tendency

## How to get a small data set into R?

Number of friends of 11 Facebook user: 
108, 103, 252, 121, 93, 57, 40, 53, 22, 116, 98.

### Assign values to a vector

```{r assign-values-to-vector}
# number of facebook friends = nff
nff <- c(108, 103, 252, 121, 93, 57, 40, 53, 22, 116, 98)
nff <- sort(nff)
nff
```

### Read data values from keyboard

After running the following code you have to set your cursor into the console and provide the data. There are 2 possibilities:
* Enter the data manually and separate each entry with >ENTER<
* Copy a string of data (e.g. from a PDF table), where each data is spearated by a blank

In both cases: Terminate the input with an extra >ENTER<

```{r read-data-from-keyboard, eval=FALSE}
nff_scan <- as.vector(scan(file = ""))
nff <- sort(nff_scan)
nff_scan
```

## The Mode

There is no mode-function in the base R module. See https://bit.ly/R-mode.
But there are many possibilites to programm this function.

Before I will demonstrate this, I need to add another number into the data set in order to get the frequencey of one number higher than the others.

```{r add-number-to-vector}
nff_mode <- c(nff, 53)
```

### Variante 1: Simple but good!

```{r compute-mode}
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
Mode(nff_mode)
```

### Version 2: with NA

```{r compute-mode-with-na}
Mode <- function(x, na.rm = FALSE) {
    if (na.rm) {
        x = x[!is.na(x)]
    }
    ux <- unique(x)
    ux[which.max(tabulate(match(x, ux)))]
}
Mode(nff_mode)
```

### Version 3: with package `modeest`

```{r compute-mode-with-modeest-pkg}
library(modeest)
mlv(nff_mode, method = "mfv")
```

## The Median

### Just the function

```{r compute-median}
median(nff)
```




### Without outlier

```{r compute-median-2}
median(nff[1:10])

```

## The Mean

### Just the function
```{r compute-mean}
mean(nff)
```

### Without outlier

```{r compute-mean-2}
mean(nff[1:10])
```

### With NA Value not removed

```{r compute-mean-3}
mean(nff[c(1:10, NA)])
```

### With NA Value removed
```{r compute-mean-4}
mean(nff[c(1:10, NA)], na.rm = TRUE)
```

### With outlier but trimmed

```{r compute-mean-5}
mean(nff, trim = 0.1)
mean(nff[2:10])
```

## The Range

### Just the function

```{r compute-range}
x <- range(nff)
xr <- x[2] - x[1]
cat("Range:", x[2], "-", x[1], "=", xr)
```

### Without outlier


```{r compute-range-2}
x <- range(nff[1:10])
xr <- x[2] - x[1]
cat("Range:", x[2], "-", x[1], "=", xr)
```

## Upper and Lower Quartile

```{r compute-upper-and-lower-quartile}
quantile(nff, type = 1)
```


## The Interquartile Range

### Just the function

```{r compute-IQR}
IQR(nff)
```

### Computed with type = 1

This measure has 9 different calculation methods (quantile algorithms) which really matter because of their big differences. Standard is type = 7 (results in 57), whereas type = 1 results in 63.

```{r compute-IQR-2}
IQR(nff, type = 1)
```

## Self-Tests

### Self-test p.25

```{r self-test-p25}
treadmill <- c(18,16,18,24,23,22,22,23,26,29,32,34,34,36,36,43,42,49,46,46,57)
Mode <- function(x, na.rm = FALSE) {
    if (na.rm) {
        x = x[!is.na(x)]
    }
    ux <- unique(x)
    tab <- tabulate(match(x, ux)); 
    ux[tab == max(tab)]
    
}
Mode(treadmill)
median(treadmill)
mean(treadmill)
quantile(treadmill, type = 6)
range(treadmill)
IQR(treadmill, type = 6)

```

### Self-Test p.27

What's the probability that someone who threw themselves of Beachy Head was 30 years or older?

* First we convert 30 into a z-score. Suppose the mean of the suicide scores was 36, and the standard deviation 13; then 30 will bekomme (30-36)/13 = `r n <- (30-36)/13; n`.
* We then look up this value in the column labelled "Bigger Portion" (i.e., the area above the value `r n`).
* I get the value of 32.28%, or put another way, there is a chance of 32.28% that a suicide victim was aged 30 or less. We can also say that there is a 67.72% chance that a suicide victim was older than 30.
