---
title: "Chapter 01 of DSUR"
author: "Peter Baumgartner"
date: "`r Sys.Date()`"
output: 
    learnr::tutorial:
        progressive: true
        allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(checkr)
knitr::opts_chunk$set(echo = FALSE)
tutorial_options(exercise.checker = checkr::check_for_learnr)
assign("my_points", 0, envir = .GlobalEnv)
```

# DSUR: Chapter 01

## Quiz

### Question 01

```{r v14-quiz-chapter-01-01}
  question("What is a confounding variable?",
    answer("A variable that is manipulated by the experimenter."),
    answer("A variable that affects the outcome being measured as well as or instead of the independent variable.", correct = TRUE),
    answer("A variable that has not been measured."),
    answer("A variable that is made up only of categories."),
    random_answer_order = TRUE,
    correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."),
    incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.")
  )
```

### Question 02


```{r quiz-chapter-01-02}
question("'Children can learn a second language faster before the age of 7'. Is this statement:",
    answer("A null hypothesis"),
    answer("A non‐scientific statement."),
    answer("A two‐tailed hypothesis."),
    answer("A one‐tailed hypothesis.", correct = TRUE)
  )
```

### Another quiz

```{r quiz1, echo=FALSE}
quiz(caption = "Quiz 1",
  question("What number is the letter A in the *English* alphabet?",
    answer("8"),
    answer("14"),
    answer("1", correct = TRUE),
    answer("23")
  ),
  question("Where are you right now? (select ALL that apply)",
    answer("Planet Earth", correct = TRUE),
    answer("Pluto"),
    answer("At a computing device", correct = TRUE),
    answer("In the Milky Way", correct = TRUE),
    incorrect = "Incorrect. You're on Earth, in the Milky Way, at a computer."
  )
)
```


## Basics: Measures of Central Tendency

## How to get a small data set into R?

Number of friends of 11 Facebook user: 
108, 103, 252, 121, 93, 57, 40, 53, 22, 116, 98.

### Assign values to a vector

```{r assign-values-to-vector}
# number of facebook friends = nff
nff <- c(108, 103, 252, 121, 93, 57, 40, 53, 22, 116, 98)
nff <- sort(nff)
nff
```

### Read data values from keyboard

After running the following code you have to set your cursor into the console and provide the data. There are 2 possibilities:
* Enter the data manually and separate each entry with >ENTER<
* Copy a string of data (e.g. from a PDF table), where each data is spearated by a blank

In both cases: Terminate the input with an extra >ENTER<

```{r read-data-from-keyboard, eval=FALSE}
nff_scan <- as.vector(scan(file = ""))
nff <- sort(nff_scan)
nff_scan
```

## The Mode

There is no mode-function in the base R module. See https://bit.ly/R-mode.
But there are many possibilites to programm this function.

Before I will demonstrate this, I need to add another number into the data set in order to get the frequencey of one number higher than the others.

```{r add-number-to-vector}
nff_mode <- c(nff, 53)
```

### Variante 1: Simple but good!

```{r compute-mode}
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
Mode(nff_mode)
```

### Version 2: with NA

```{r compute-mode-with-na}
Mode <- function(x, na.rm = FALSE) {
    if (na.rm) {
        x = x[!is.na(x)]
    }
    ux <- unique(x)
    ux[which.max(tabulate(match(x, ux)))]
}
Mode(nff_mode)
```

### Version 3: with package `modeest`

```{r compute-mode-with-modeest-pkg}
library(modeest)
mlv(nff_mode, method = "mfv")
```

## The Median

### Just the function

```{r compute-median}
median(nff)
```




### Without outlier

```{r compute-median-2}
median(nff[1:10])

```

## The Mean

### Just the function
```{r compute-mean}
mean(nff)
```

### Without outlier

```{r compute-mean-2}
mean(nff[1:10])
```

### With NA Value not removed

```{r compute-mean-3}
mean(nff[c(1:10, NA)])
```

### With NA Value removed
```{r compute-mean-4}
mean(nff[c(1:10, NA)], na.rm = TRUE)
```

### With outlier but trimmed

```{r compute-mean-5}
mean(nff, trim = 0.1)
mean(nff[2:10])
```

## The Range

### Just the function

```{r compute-range}
x <- range(nff)
xr <- x[2] - x[1]
cat("Range:", x[2], "-", x[1], "=", xr)
```

### Without outlier


```{r compute-range-2}
x <- range(nff[1:10])
xr <- x[2] - x[1]
cat("Range:", x[2], "-", x[1], "=", xr)
```

## Upper and Lower Quartile

```{r compute-upper-and-lower-quartile}
quantile(nff, type = 1)
```


## The Interquartile Range

### Just the function

```{r compute-IQR}
IQR(nff)
```

### Computed with type = 1

This measure has 9 different calculation methods (quantile algorithms) which really matter because of their big differences. Standard is type = 7 (results in 57), whereas type = 1 results in 63.

```{r compute-IQR-2}
IQR(nff, type = 1)
```

## Self-Tests

### Self-test p.25

```{r self-test-p25}
treadmill <- c(18,16,18,24,23,22,22,23,26,29,32,34,34,36,36,43,42,49,46,46,57)
Mode <- function(x, na.rm = FALSE) {
    if (na.rm) {
        x = x[!is.na(x)]
    }
    ux <- unique(x)
    tab <- tabulate(match(x, ux)); 
    ux[tab == max(tab)]
    
}
Mode(treadmill)
median(treadmill)
mean(treadmill)
quantile(treadmill, type = 6)
range(treadmill)
IQR(treadmill, type = 6)

```

### Self-Test p.27

What's the probability that someone who threw themselves of Beachy Head was 30 years or older?

* First we convert 30 into a z-score. Suppose the mean of the suicide scores was 36, and the standard deviation 13; then 30 will bekomme (30-36)/13 = `r n <- (30-36)/13; n`.
* We then look up this value in the column labelled "Bigger Portion" (i.e., the area above the value `r n`).
* I get the value of 32.28%, or put another way, there is a chance of 32.28% that a suicide victim was aged 30 or less. We can also say that there is a 67.72% chance that a suicide victim was older than 30.
