---
title: "DSUR Chap. 01"
author: "Peter Baumgartner"
date: "`r Sys.Date()`"
output: 
    learnr::tutorial:
        progressive: true
        allow_skip: true
        theme: cerulean
        css: css/mystyle.css
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
library(checkr)
library(ggplot2)
library(tidyverse)
library(modeest)
tutorial_options(exercise.checker = checkr::check_for_learnr)
knitr::opts_chunk$set(
        message = T,
        error = T,
        warning = T,
        comment = '##',
        highlight = T,
        prompt = F,
        echo = F
        )

# Global Variables --------------------------------------------------------------
source('www/check_test.R')
theme_update(plot.title = element_text(hjust = 0.5, face = "bold"))
nff <- c(108, 103, 252, 121, 93, 57, 40, 53, 22, 116, 98)
```

# DSUR: Chapter 01

## Overview of Chapter 01

> In essence, I assume you know nothing about statistics, but I do assume you have some very basic grasp of computers ... and math. (DSUR, p.XXV)



### Basic notions

The first chapter (pp 1--31) covers some basic concepts in a general way, e.g. without using R. The examples uses small data sets and are easy to calculate by hand. But nevertheless I have tried to find a pedant in R. This was also a kind of exercise for me to get familiar with basic statistic concepts in R.

This chapters could be divided into two part: It beginns with an introduction to the research process: How to start with initial observations, how to generate theories and hypothesis to test the assumptions?

Then many basic notions are explained:

+ Independent and dependent variables
+ Levels of measurement
+ Measurement error
+ Validity and reliability
+ Correlation and experimental research methods
+ Randomization

To check if you are conversant with these notions you could take the [multiple choice test for this chapter](https://studysites.uk.sagepub.com/dsur/study/mcq/Chapter%201/quiz.htm) taken from the student resources of the companion site. Additionally I have converted another [13 questions](#quiz01) from the part of the instructor resources.

### Basic information for analysing data

The second part of the chapter provides basic information for analysing data:

+ Frequency distribution
+ Measures of the central tendency
    - The mode
    - The median
    - The mean
+ Dispersion measures
    - The range
    - The quartiles
+ Going beyond data with a frequency distribution
    - Probability distribution
    - Z-scores
+ Fitting statistical models to data
    - Null hypothesis
    - Alternative (experimental) hypthesis

### Summary

Recapitulate in your mind the following questions and compare it with the text in the book chapter:

+ How do I do research?
+ What qualities should a scientific theory have?
+ What is the difference between reliability and validity?
+ What is the difference between experimental and correlation research?
+ Why is randomization important?
+ What is a frequency distribution and when is it normal?
+ What are the mode, median and mean?
+ What is the normal distribution?

## Frequency distributions

### Introductory video

<!-- <p><img src="https://www.youtube.com/watch?v=24YCewQ8s3M" /></p> -->

## Explore skewed distrubtions

```{r my-dist-ui1, context="render"}
    fluidRow(
        column(3,
            radioButtons("type", "Choose a distribution:",
                           c("Normal" = "norm",
                             "Right-skewed" = "right",
                             "Left-skewed" = "left")
                        )
            ),
        column(9, htmlOutput("distText"))
    ) # end of fluidRow
    fluidRow(
        column(8,
              plotOutput("distPlot")
              )
)





```



```{r my-dist-server1,  context="server"}

explainNormal <- "This is the <strong>normal</strong> distribution"
explainRight  <- "This is the right-skewed distribution"
explainLeft   <- "This is the left-skewed distribution"

 output$distPlot <- renderPlot({
    switch(input$type,
        norm    = {histo <- as_tibble(rbeta(10000,5,5))
                  caption <- "Normal distribution"
                  output$distText <- renderText(explainNormal)},
        right   = {histo <- as_tibble(rbeta(10000,2,5))
                  caption <- "Positivey or right-skewed distribution"
                  output$distText <- renderText(explainRight)},
        left    = {histo <- as_tibble(rbeta(10000,5,2))
                  caption <- "Negatively or left-skewed distribution"
                  output$distText <- renderText(explainLeft)}
      )
ggplot(histo, aes(value)) +
        geom_histogram(aes(y = ..density..),
                       alpha = 0.25, fill = "blue", binwidth = .025) +
        stat_function(fun = dnorm,
                      args = c(mean = mean(histo$value),
                               sd = sd(histo$value)),
                      col = "tomato") +
        labs(
            title = caption,
            x = "Score",
            y = "Density"
        ) +
    labs(caption = "lines: black = mode, yellow = mean, green = median")
})
```

## Frequency distributions: Kurtosos

### Introductory video

### Explore

## The Mode

### Personal remark

When I first looked how to calculate the mode of a vector I was suprised to learn that there is no mode function in the `base` R module. Or better said, there is a `mode()` function, but it tells you the internal storage mode of the object, not the value that occurs the most in its argument.

According to the interesting, but high-level discussion at [StackOverflow](https://stackoverflow.com/questions/2547402/is-there-a-built-in-function-for-finding-the-mode) there are several possibilities.


I will pick two of them:

+ Using the `modeest` package.
+ Programming a `Mode()` function


### Preparing the data set

> The mode is the most frequently occuring score of a data set.

Before I will demonstrate how to to calculate the mode in R, I need to add at least another number into our `nff` data set in order to get the frequencey of (at least) one number higher than the others. In the following example I will add _two_ numbers to simulate a bimodal distribution as some programming solutions return only the first detected mode.

As we will use this extra numbers just for the mode demonstration of this section, the following chunk creates a new variable `nff_mode` just for this purpose.

```{r prepare-nff-mode}
nff_mode <- c(nff, 53, 108)
nff_mode <- sort(nff_mode)
```


```{r add-number-to-vector, exercise=TRUE, exercise.setup = "prepare-nff-mode", exercise.startover=FALSE}
nff
nff_mode <- c(nff, 53, 108)
nff_mode <- sort(nff_mode)
nff_mode
```

Run the above chunk of code to inspect the new data set. Note that the first vector `nff` is two items shorter and that `nff_mode` has two most frequently values: 53 and 108. Then go to the next subsection.

### Use of the `modeest` package

The three lines of the next R chunk are explained as follows:

1. The first line `library(modeest)` loads the package into memory. If you are working without the interactive tutorial you have probably to install the package with `install.packages("modeest")`. You will learn how to install and load packages later in another chapter of this tutorial. Here I just want to mention that a huge community of nice guys have today (2018-06-11) already 12.618 packages developed and uploaded for free use. These many packages provide additional functionality to the R basis modules.

2. The second line is just to remind you of the content of our vector `nff_mode`.

3. `mfv()` is one of the many functions of the `modeest`package. It returns the **m**ost **f**requent **v**alue(s) in a given numerical vector. We use this function with our vector variable.

```{r compute-mode-with-mfv, message = FALSE, exercise=TRUE, exercise.setup = "prepare-nff-mode", exercise.startover=FALSE}
library(modeest)
nff_mode
mfv(nff_mode)
```


### Write your own Mode function

As programming R functions is an advanced topic I will just provide some information in this section and abstain from interactive exercises. You are free to skip this section completly and continue with [The Median](#section-the-median).

The following R chunk code looks a little bit frightening So let's try it out and then I explain how it works in detail.


```{r write-Mode-function, exercise=TRUE, exercise.setup = "prepare-nff-mode", exercise.startover=FALSE}
Mode <- function(x) {
    ux <- unique(x)
    tab <- tabulate(match(x, ux))
    ux[tab == max(tab)]
}
Mode(nff_mode)
```

Execept of the last line the above code chunk generates a function with the name Mode. The capital 'M' in our `Mode()` function is important as R is case sensitive. With a small 'm' it would return the storage mode of an object and not the most frequent value(s).

The parenthesis `(x)` signals that we have to deliver to this function a value. In our case -- this is the last line of the chunk -- we call the Mode function with the vector variable `nff_mode`.

All commands inside the function have to be written between curly braces `{}`. The last line before the closing curly brace returns a value to the calling function. We could have used explicitly the `return()` function, but this is in R not common.

We could have captured the result of the function by assigning it to a new variable like `t <- Mode(nff_mode)`. But this is here not necessary.

The above programmed new 'Mode'-function uses `unique()`. The unique function returns a vector, with duplicate elements removed.

The next line uses two functions and stores the result:

1. `match()` returns a vector of the positions of matches of its first argument in its second. The elements of our vector `nff_mode`  (which has gotten the name `x` inside the function) are matched with the unique values stores in the variable `ux`.

2. The match function itself is applied by `tabulate()`, which takes the integer-valued vector and counts the number of times each integer occurs in it.

3. The value of this computation is assigned to a new variable `tab`.

The last line uses the result of the previous calculation (stored in `tab`) to subset (= `[]`) all maxima, which themselves are returned by the `max()` function.

This is quite a mouthful of a combination of different functions. But keep in mind that this explanation is just for demonstration purposes. It is much easier to use `mfv()` of the `modeest` package. -- Or to inspect the graph of the frequency distribution in order to detect the mode(s)…

***
Note: To work with missing values (so called `NA`s, = data **n**ot **a**vailable') we would need to modify our `Mode()` function as exemplified in the [answer by jprockbelly](https://stackoverflow.com/a/25635740/7322615).



## The Median

### Just the function

In contrast to mode there is a special function `median()`. Write the command to compute the median for the `nff` vector variable.

```{r compute-the-median, exercise=TRUE}

```



```{r compute-the-median-check}
check_compute_median(USER_CODE)
```



### Without outlier

```{r compute-median-2}
median(nff[1:10])

```

## The Mean

### Just the function
```{r compute-mean}
mean(nff)
```

### Without outlier

```{r compute-mean-2}
mean(nff[1:10])
```

### With NA Value not removed

```{r compute-mean-3}
mean(nff[c(1:10, NA)])
```

### With NA Value removed
```{r compute-mean-4}
mean(nff[c(1:10, NA)], na.rm = TRUE)
```

### With outlier but trimmed

```{r compute-mean-5}
mean(nff, trim = 0.1)
mean(nff[2:10])
```

## The Range

### Just the function

```{r compute-range}
x <- range(nff)
xr <- x[2] - x[1]
cat("Range:", x[2], "-", x[1], "=", xr)
```

### Without outlier


```{r compute-range-2}
x <- range(nff[1:10])
xr <- x[2] - x[1]
cat("Range:", x[2], "-", x[1], "=", xr)
```

## Upper and Lower Quartile

```{r compute-upper-and-lower-quartile}
quantile(nff, type = 1)
```


## The Interquartile Range

### Just the function

```{r compute-IQR}
IQR(nff)
```

### Computed with type = 1

This measure has 9 different calculation methods (quantile algorithms) which really matter because of their big differences. Standard is type = 7 (results in 57), whereas type = 1 results in 63.

```{r compute-IQR-2}
IQR(nff, type = 1)
```

## Self-Tests

### Self-test p.25

```{r self-test-p25}
treadmill <- c(18,16,18,24,23,22,22,23,26,29,32,34,34,36,36,43,42,49,46,46,57)

Mode(treadmill)
median(treadmill)
mean(treadmill)
quantile(treadmill, type = 6)
range(treadmill)
IQR(treadmill, type = 6)

```

### Self-Test p.27

What's the probability that someone who threw themselves of Beachy Head was 30 years or older?

* First we convert 30 into a z-score. Suppose the mean of the suicide scores was 36, and the standard deviation 13; then 30 will bekomme (30-36)/13 = `r n <- (30-36)/13; n`.
* We then look up this value in the column labelled "Bigger Portion" (i.e., the area above the value `r n`).
* I get the value of 32.28%, or put another way, there is a chance of 32.28% that a suicide victim was aged 30 or less. We can also say that there is a 67.72% chance that a suicide victim was older than 30.

## Quiz for Chapter 01 {#quiz01}

<!-- ### Question 01 -->

<!-- ```{r quiz-chapter-01-01} -->
<!--   question("What is a confounding variable?", -->
<!--     answer("A variable that is manipulated by the experimenter."), -->
<!--     answer("A variable that affects the outcome being measured as well as or instead of the independent variable.", correct = TRUE), -->
<!--     answer("A variable that has not been measured."), -->
<!--     answer("A variable that is made up only of categories."), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!--   ) -->
<!-- ``` -->

<!-- ### Question 02 -->


<!-- ```{r quiz-chapter-01-02} -->
<!-- question("'Children can learn a second language faster before the age of 7'. Is this statement:", -->
<!--     answer("A null hypothesis"), -->
<!--     answer("A non‐scientific statement."), -->
<!--     answer("A two‐tailed hypothesis."), -->
<!--     answer("A one‐tailed hypothesis.", correct = TRUE), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!--   ) -->
<!-- ``` -->

<!-- ### Question 03 -->

<!-- ```{r quiz-chapter-01-03} -->
<!-- question("If a psychological test is valid, what does this mean?", -->
<!--     answer("The test will give consistent results."), -->
<!--     answer("The test measures what it claims to measure.", correct = TRUE), -->
<!--     answer("The test has internal consistency."), -->
<!--     answer("The test measures a psychologically useful variable."), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 04 -->

<!-- ```{r quiz-chapter-01-04} -->
<!-- question("If my null hypothesis is ‘Dutch people do not differ from English people in height’, what is my alternative hypothesis?", -->
<!--     answer("Dutch people are taller than English people."), -->
<!--     answer("English people are taller than Dutch people."), -->
<!--     answer("Dutch people differ in height from English people."), -->
<!--     answer("All of the above are plausible alternative hypotheses.", correct = TRUE), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 05 -->

<!-- ```{r quiz-chapter-01-05} -->
<!-- question("When questionnaire scores predict, or correspond with, external measures of the same construct that the questionnaire measures it is said to have:", -->
<!--     answer("Ecological validity."), -->
<!--     answer("Factorial validity."), -->
<!--     answer("Content validity."), -->
<!--     answer("Criterion validity.", correct = TRUE), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 06 -->

<!-- ```{r quiz-chapter-01-06} -->
<!-- question("A variable manipulated by a researcher is known as:", -->
<!--     answer("A dependent variable."), -->
<!--     answer("A confounding variable."), -->
<!--     answer("A discrete variable."), -->
<!--     answer("An independent variable.", correct = TRUE), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 07 -->

<!-- ```{r quiz-chapter-01-07} -->
<!-- question("A predictor variable is another name for:", -->
<!--     answer("A dependent variable."), -->
<!--     answer("A confounding variable."), -->
<!--     answer("A discrete variable."), -->
<!--     answer("An independent variable.", correct = TRUE), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 08 -->

<!-- ```{r quiz-chapter-01-08} -->
<!-- question("What kind of variable is IQ, measured by a standard IQ test?", -->
<!--     answer("Categorical."), -->
<!--     answer("Discrete."), -->
<!--     answer("Nominal."), -->
<!--     answer("Continuous.", correct = TRUE), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 09 -->

<!-- ```{r quiz-chapter-01-09} -->
<!-- question("A frequency distribution in which high scores are most frequent (i.e. bars on the graph are highest on the right‐hand side) is said to be:", -->
<!--     answer("Positively skewed."), -->
<!--     answer("Leptokurtic."), -->
<!--     answer("Platykurtic."), -->
<!--     answer("Negatively skewed.", correct = TRUE), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 10 -->

<!-- ```{r quiz-chapter-01-10} -->
<!-- question("A frequency distribution in which there are too few scores at the extremes of the distribution is said to be:", -->
<!--     answer("Positively skewed."), -->
<!--     answer("Leptokurtic.", correct = TRUE), -->
<!--     answer("Platykurtic."), -->
<!--     answer("Negatively skewed."), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 11 -->

<!-- ```{r quiz-chapter-01-11} -->
<!-- question("Which of the following is designed to compensate for practice effects?", -->
<!--     answer("A repeated measures design."), -->
<!--     answer("Randomization of participants."), -->
<!--     answer("Counterbalancing.", correct = TRUE), -->
<!--     answer("A control condition."), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 12 -->

<!-- ```{r quiz-chapter-01-12} -->
<!-- question("Variation due to variables that have not been measured is known as:", -->
<!--     answer("Unsystematic variance.", correct = TRUE), -->
<!--     answer("Homogeneous variance."), -->
<!--     answer("Systematic variance."), -->
<!--     answer("Model variance."), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->

<!-- ### Question 13 -->

<!-- ```{r quiz-chapter-01-13} -->
<!-- question("If the scores on a test have a mean of 26 and a standard deviation of 4, what is the z‐score for a score of 18?", -->
<!--     answer("-2", correct = TRUE), -->
<!--     answer("11"), -->
<!--     answer("2"), -->
<!--     answer("-1.41"), -->
<!--     random_answer_order = TRUE, -->
<!--     correct = paste("Correct: You've reached", assign("my_points", my_points + 1, envir = .GlobalEnv), " of 13 possible points."), -->
<!--     incorrect = paste("Incorrect: You've reached", assign("my_points", my_points - 1, envir = .GlobalEnv), " of 13 possible points.") -->
<!-- ) -->
<!-- ``` -->



### 13 Questions

To be used in conjunction with Field, A. P., Miles, J., and Field, Z. C. (2012). [Discovering statistics using R](https://studysites.uk.sagepub.com/dsur/main.htm). London: Sage

```{r quiz-chapter-01}

quiz(caption = "Quiz Chapter 01",
  question("What is a confounding variable?",
    answer("A variable that is manipulated by the experimenter."),
    answer("A variable that affects the outcome being measured as well as or instead of the independent variable.", correct = TRUE),
    answer("A variable that has not been measured."),
    answer("A variable that is made up only of categories."),
    random_answer_order = TRUE
  ),
question("'Children can learn a second language faster before the age of 7'. Is this statement:",
    answer("A null hypothesis"),
    answer("A non‐scientific statement."),
    answer("A two‐tailed hypothesis."),
    answer("A one‐tailed hypothesis.", correct = TRUE),
    random_answer_order = TRUE
  ),

question("If a psychological test is valid, what does this mean?",
    answer("The test will give consistent results."),
    answer("The test measures what it claims to measure.", correct = TRUE),
    answer("The test has internal consistency."),
    answer("The test measures a psychologically useful variable."),
    random_answer_order = TRUE
),

question("If my null hypothesis is ‘Dutch people do not differ from English people in height’, what is my alternative hypothesis?",
    answer("Dutch people are taller than English people."),
    answer("English people are taller than Dutch people."),
    answer("Dutch people differ in height from English people."),
    answer("All of the above are plausible alternative hypotheses.", correct = TRUE),
    random_answer_order = TRUE
),

question("When questionnaire scores predict, or correspond with, external measures of the same construct that the questionnaire measures it is said to have:",
    answer("Ecological validity."),
    answer("Factorial validity."),
    answer("Content validity."),
    answer("Criterion validity.", correct = TRUE),
    random_answer_order = TRUE
),

question("A variable manipulated by a researcher is known as:",
    answer("A dependent variable."),
    answer("A confounding variable."),
    answer("A discrete variable."),
    answer("An independent variable.", correct = TRUE),
    random_answer_order = TRUE
),

question("A predictor variable is another name for:",
    answer("A dependent variable."),
    answer("A confounding variable."),
    answer("A discrete variable."),
    answer("An independent variable.", correct = TRUE),
    random_answer_order = TRUE
),

question("What kind of variable is IQ, measured by a standard IQ test?",
    answer("Categorical."),
    answer("Discrete."),
    answer("Nominal."),
    answer("Continuous.", correct = TRUE),
    random_answer_order = TRUE
),

question("A frequency distribution in which high scores are most frequent (i.e. bars on the graph are highest on the right‐hand side) is said to be:",
    answer("Positively skewed."),
    answer("Leptokurtic."),
    answer("Platykurtic."),
    answer("Negatively skewed.", correct = TRUE),
    random_answer_order = TRUE
),

question("A frequency distribution in which there are too few scores at the extremes of the distribution is said to be:",
    answer("Positively skewed."),
    answer("Leptokurtic.", correct = TRUE),
    answer("Platykurtic."),
    answer("Negatively skewed."),
    random_answer_order = TRUE
),

question("Which of the following is designed to compensate for practice effects?",
    answer("A repeated measures design."),
    answer("Randomization of participants."),
    answer("Counterbalancing.", correct = TRUE),
    answer("A control condition."),
    random_answer_order = TRUE
),

question("Variation due to variables that have not been measured is known as:",
    answer("Unsystematic variance.", correct = TRUE),
    answer("Homogeneous variance."),
    answer("Systematic variance."),
    answer("Model variance."),
    random_answer_order = TRUE
),

question("If the scores on a test have a mean of 26 and a standard deviation of 4, what is the z‐score for a score of 18?",
    answer("-2", correct = TRUE),
    answer("11"),
    answer("2"),
    answer("-1.41"),
    random_answer_order = TRUE
    )
)
```


<!-- ## Measures of Central Tendency -->

