---
title: "DSUR Tutorial"
author: "Peter Baumgartner"
date: "`r Sys.Date()`"
output: 
    learnr::tutorial:
        progressive: true
        allow_skip: true
        theme: cerulean
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
library(checkr)
tutorial_options(exercise.checker = checkr::check_for_learnr)
knitr::opts_chunk$set(
        message = T,
        error = T,
        warning = T,
        comment = '##',
        highlight = T,
        prompt = F,
        echo = F
        )

```

```{r}
# this chunk (or line) is necessary,
# because there is at the moment no interactive exercise
# without this chunk (or line) learnr generates the following error:
# "Warning: Error in download$filter: attempt to apply non-function
#   [No stack trace available]"
learnr::initialize_tutorial()
```



# Introduction

## Material and subjects covered

This tutorial is supplementary material for [Discovering statistics using R](https://studysites.uk.sagepub.com/dsur/main.htm) by Field, A. P., Miles, J., and Field, Z. C. (2012, London: Sage). In this tutorial I provide some interactive exercises and quizzes to learn statistics **and** R. 

I used 'and' instead of 'with' or 'using' as most of the interactive exercises aks to apply R code for a given statistical procedure. So learning R is more prominent as in the mentioned book, whereas the statistical content relies heavily on the explanations given in `Discovering statistics using R` (abbreviated with `DSUR`). To benefit most of this course material you definitively have to buy a copy of DSUR.


<center>

[
![Field, A., Miles, J., & Field, Z. (2012). Discovering Statistics Using R (1st ed.). SAGE Publications Ltd.
](https://images-na.ssl-images-amazon.com/images/I/41M2PgYiZXL._SX366_BO1,204,203,200_.jpg)](https://studysites.uk.sagepub.com/dsur/main.htm)

</center>

## Table of contents

00-Chapter: Introduction (this file)   
[01-Chapter](https://pbaumgartner.shinyapps.io/DSURT-01/): Some basic concepts and measures   
02-Chapter: A gentle introduction into statistics   
03-Chapter: The R and RStudio environment   
04-Chapter: Exploring data with graphs   
05-Chapter: Exploring assumptions   
06-Chapter: Correlation   
07-Chapter: Regression   
08-Chapter: Logistic Regression   
09-Chapter: Comparing two means   
10-Chapter: Comparing several means: ANOVA (GLM 1)  
11-Chapter: Analysis of covariance: ANCOVA (GLM 2)    
12-Chapter: Factorial ANOVA (GLM 3)  
13-Chapter: Repeated-measures design (GLM 4)   
14-Chapter: Mixed designs (GLM 5)   
15-Chapter: Non-parametric tests   
16-Chapter: Multivariate analysis of variance (MANOVA)   
17-Chapter: Exploratory factor analysis   
18-Chapter: Categorial data   
19-Chapter: Multilevel linear models 

## Target group and prerequisite 

This interactive tutorial is primarily intended for people who work through DSUR and want to have more interactive exercises using R. I assume the same level of background knowledge as in DSUR, namely nothing:

"_In essence, I assume you know nothing about statistics, but I do assume you have some very basic grasp of computers … and math._" (DSUR, p.XXV)

## Learning outcomes

The following learning outcomes provide information at the course level. As there are different levels in a curriculum (course, module/chapter lesson/section etc.) with their own learning outcomes only 3-5 learning outcomes on each level is sufficent. In formulating the following outcomes I am applying the taxonomy by Anderson & Krathwohl (2001, 2002) which is an extended and revised edition of the original taxonomy by Bloom (1956).

Having successfully completed this interactive course you will be able to:

+ manipulate and transform data with R to bring them in a format suitable for exploratory data analysis
+ create high quality data visualizations with the R package ggplot2
+ implementing of statistical procedures within the R environment
+ programming an analysis with R in such a way that the script can be used with minimal effort for similar datasets and analyses

**References:**

+ Anderson, L. W. (2002). Revising Bloom’s taxonomy. Columbus, OH: Ohio State University.
+ Anderson, L. W., & Krathwohl, D. R. (2001). A taxonomy for learning, teaching, and assessing: a revision of Bloom’s taxonomy of educational objectives. New York: Longman.
+ Bloom, B. S., & Krathwohl, D. R. (1956). Taxonomy of Educational Objectives Book 1: Cognitive Domain (2nd ed.). Longman.


## Structure of this tutorial {#structure}

I tried to follow as exactly as possible the structure of DSUR. But sometimes I had to develop other sequeneces. This is esepcially true for the first two chapters as DSUR starts with explaining R in the third chapter. To provide for the R newbies some meaningful exercises even for the first two chapters I had to antedate some content (e.g. how to install R, to explain the elements of the graphical user interface `GUI` etc.

Sometimes I added R related information or even decided to depart from DSUR content. The last is especially true with `RStudio` which installation and GIUs I used instead of the standard R Interface or the program package `R Commander`. But I have noted this differences within a yellow box.

<div class="alert alert-warning" role="alert">
  In yellow boxes I will mention differences to the accompanying source of DSUR. 
</div>

## How to use this tutorial?

[Here I will explain the different display modes and required actions]

## Learning Outcome {#learning-outcome}

## Tools used and source code

This interactive course is written in R Markdown and uses the [`learnr`](https://rstudio.github.io/learnr/) package in combination with the [`checkr`](https://github.com/dtkaplan/checkr). The source code can be inspected at [my GitHub account](https://github.com/petzi53/DSUR). 

I have written this tutorial to rebuild my own competences in statistics and at the same time to improve my R practice. As I am not an expert on both subject please help to improve this tutorial via PRs or [contact me](mailto:peter_baumgartner@me.com) for errors or other improvements.

